The \dealii{} Library, Version 8.5

### 2017-03-04 8.5/paper.tex
Extend text

@@

For the new release, the geometric multigrid facilities in \dealii{} have been
thoroughly overhauled regarding their scalability on large-scale parallel
computers. During this process, a geometric multigrid implementation based on
the fast matrix-free kernels from \cite{KronbichlerKormann2012} has been
benchmarked up to 147,456 cores. The fast matrix-vector product revealed
several scalability bottlenecks, including unnecessary inner products inside
the Chebyshev smoother and $\mathcal O(n_\text{levels})$ global communication
steps during the restriction process rather than only the single global
communication step that is necessary when going to the coarest grid. New
matrix-free transfer implementations called \texttt{MGTransferMatrixFree} were
devised that can replace the matrix-based \texttt{MGTransferPrebuilt} class
for tensor product elements. Besides better scalability than the Trilinos
Epetra matrices underlying the latter, the matrix-free transfer is also a much
better for high-order elements with complexity per degree of freedom of
$\mathcal O(d p)$ in the polynomial degree $p$ in $d$ dimensions rather than
$\mathcal O(p^d)$ for the matrix-based approach.

@@

  \caption{Scaling of multigrid algorithms on SuperMUC.}

@@

The scalability of the improved geometric multigrid framework is shown in
Fig.~\ref{fig:scaling_mg}, including a combined strong and weak scaling plot
in the left panel using continuous $\mathcal Q_3$ elements with 57~million to
232~billion degrees of freedom for discretizing the Laplacian. Along each
line, the same problem size is solved with an increasing number of cores,
whereas different lines are a factor of eight apart and always start at
3.5~million degrees of freedom per core. Almost ideal scalability down to
approximately 0.1~seconds can be observed also on 147k~cores. The right panel
of Fig.~\ref{fig:scaling_mg} shows the effect of the aforementioned
algorithmic improvements on a setup with discontinuous DG elements, clearly
improving the latency of the multigrid V-cycle. The improved algorithms are
shown in the updated step-37 tutorial program.

@@

\verb!FE_Enriched! finite element implements a partition of unity finite element method (PUM) by Babuska and Melenk which enriches a standard finite element with an enrichment function multiplied with another (usually linear) finite element. This allows to include in the finite element space a priori knowledge about the partial differential equation being solved which in turn improves the local approximation properties of the spaces. The user can also use enriched and non-enriched finite elements in different parts of the domain. 
\verb|!DoFTools::make_hanging_node_constraints()| function can automatically make the resulting space $C^0$ continuous.
Existing \verb|SolutionTransfer| class can be used to transfer the solution during $h$\,-adaptive refinement from a coarse to a fine mesh under the condition that all child elements are also enriched.

@@

Facilitate the usage of matrix-free method by providing a \verb!MatrixFreeOperator::Base! class,
which implements various interface to matrix-vector products, including necessary operations when used in
the context of the geometric multigrids, methods needed for usage within the linear operator as well as with Jacobi preconditioner.
The derived classes only need to implement \verb!apply_add()! method that is
used in the \verb!vmult()! functions, and a method to compute the diagonal entries of the underlying matrix.
The \verb!MatrixFreeOperator! namespace contains implementations of \verb!MatrixFreeOperators::LaplaceOperator! and
\verb!MatrixFreeOperators::MassOperator!.

The framework was also included in the updated step-37 tutorial program.

@@

Switch default of Lagrange elements to Gauss-Lobatto

>>>>

\item Various improvements for high-order elements, including a switch of
  support points in \texttt{FE\_Q} and \texttt{FE\_DGQ} to Gauss--Lobatto
  support points, more stable evaluation of Legendre polynomials, and several
  bugfixes for high-order polynomial mappings defined through the
  \texttt{MappingQ} class.

@@

For the new release, the geometric multigrid facilities in \dealii{} have been
thoroughly overhauled regarding their scalability on large-scale parallel
systems. During this process, a geometric multigrid implementation based on
the fast matrix-free kernels from \cite{KronbichlerKormann2012} has been
benchmarked on up to 147,456 cores. The fast matrix-vector products revealed
several scalability bottlenecks in the other multigrid components, including
unnecessary inner products inside the Chebyshev smoother and
$\mathcal O(n_\text{levels})$ global communication steps during the
restriction process rather than the single unavoidable global communication
step inherent to going to the coarsest grid and the coarse solve. New
matrix-free transfer implementations called \texttt{MGTransferMatrixFree} were
devised that can replace the matrix-based \texttt{MGTransferPrebuilt} class
for tensor product elements. Besides better scalability than the Trilinos
Epetra matrices underlying the latter, the matrix-free transfer is also a much
better for high-order elements with a complexity per degree of freedom of
$\mathcal O(d p)$ in the polynomial degree $p$ in $d$ dimensions rather than
$\mathcal O(p^d)$ for the matrices.

@@

  \caption{Scaling of \dealii{}'s geometric multigrid algorithms on SuperMUC.}

@@

The scalability of the improved geometric multigrid framework is shown in
Fig.~\ref{fig:scaling_mg}, including a combined strong and weak scaling plot
in the left panel using continuous $\mathcal Q_3$ elements with 57~million to
232~billion degrees of freedom for discretizing the Laplacian. Along each
line, the same problem size is solved with an increasing number of cores,
whereas different lines are a factor of eight apart and always start at
3.5~million degrees of freedom per core with an absolute performance of around
650,000 degrees of freedom per core and second. Almost ideal scalability down
to approximately 0.1~seconds can be observed also on 147k~cores. The right
panel of Fig.~\ref{fig:scaling_mg} shows the effect of the aforementioned
algorithmic improvements on a setup with discontinuous DG elements, clearly
improving the latency of the multigrid V-cycle. The updated step-37 tutorial
programs presents the updated algorithms and the MPI-parallel multigrid
setting with matrix-free operator evaluation.

@@

\verb!FE_Enriched! finite element implements a partition of unity finite element method (PUM) by Babuska and Melenk which enriches a standard finite element with an enrichment function multiplied with another (usually linear) finite element. This allows to include in the finite element space a priori knowledge about the partial differential equation being solved which in turn improves the local approximation properties of the spaces. The user can also use enriched and non-enriched finite elements in different parts of the domain.
\verb|DoFTools::make_hanging_node_constraints()| function can automatically make the resulting space $C^0$ continuous.
Existing \verb|SolutionTransfer| class can be used to transfer the solution during $h$\,-adaptive refinement from a coarse to a fine mesh under the condition that all child elements are also enriched.

@@

In order to facilitate the usage of matrix-free methods, a \verb!MatrixFreeOperator::Base! class has been introduced,
implementing various interfaces to matrix-vector products and the necessary operations for the interface residuals according to \cite{JanssenKanschat2011} in
the context of the geometric multigrids. Furthermore, the class is compatible with the linear operator framework and provides an interface to a Jacobi preconditioner.
The derived classes only need to implement the \verb!apply_add()! method that is
used in the \verb!vmult()! functions, and a method to compute the diagonal entries of the underlying matrix.
The \verb!MatrixFreeOperator! namespace contains implementations of \verb!MatrixFreeOperators::LaplaceOperator! and
\verb!MatrixFreeOperators::MassOperator!.
The updated step-37 tutorial program makes use of these facilities and
explains their usage in detail. Using the matrix-free mass operator,
\verb!VectorTools::project! has become much faster than the previous
matrix-based approach and also works in parallel with MPI.

@@

High-order Lagrange elements, both continuous \verb!FE_Q! and discontinuous
\verb!FE_DGQ! types, now use the nodal points of the Gauss--Lobatto
quadrature formula as support points, rather than the previous equidistant
ones. For cubic polynomials and higher, the point distribution has changed and
thus the entries in solution vectors will look different now. Note, however,
that using the Gauss--Lobatto points as nodal points results in much more
stable interpolation, including better iteration counts in most iterative
solvers.
